Pulmonary acini exhibit complex changes during postnatal rat lung development

### 2021-06-08 manuscript.tex
Add additional citation, suggested by Reviewer1 [ci skip]
This build is based on
https://github.com/habi/acinar-analysis-manuscript/commit/46c4c67bf996034873f7ab33969af2db0093951a.

This commit was created by the following CI build and job:
https://github.com/habi/acinar-analysis-manuscript/commit/46c4c67bf996034873f7ab33969af2db0093951a/checks
https://github.com/habi/acinar-analysis-manuscript/runs/918798125

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681}.
It has been submited to \href{https://journals.plos.org/plosone/}{\emph{PLOS ONE}}.
The manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/5e778036e316c47b8bac4ce50086b191a7e3707a...f0b634e60b036c18080d89c63034379371acaf83}{has been updated} since submission.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/f0b634e60b036c18080d89c63034379371acaf83/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/f0b634e60b036c18080d89c63034379371acaf83}{habi/acinar-analysis-manuscript@f0b634e}
on June 8, 2021.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(f0b634e)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(f0b634e)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\end{itemize}

@@

As shown in a study by Osmanagic et al.~{[}\protect\hyperlink{ref-LzalnLBH}{49}{]} the total alveolar number shows a large variability between different labs.
E.g.~the reported total number of alveoli in adult C57BL6 mice differs by a factor of approximately 8.
All the studies mentioned by Osmanagic et al.~were performed in labs well-known for their quality of stereology-based investigations of lung morphology.
Therefore, any simple bias or methodical error may not account for the reported differences.
It seems to be that the total number of alveoli is dependent on unknown small methodical differences and minute differences in how the operator actually does count.
The seemingly minuscule volume fraction of lung studied (around 1 \% or less) in this manuscript is---according to the stereological guidelines {[}\protect\hyperlink{ref-dNc8FfNn}{28}{]}---ample volume to gain mathematically correct results if the stereology is carried out correctly as we have done to the best of our knowledge.
For human lungs for example a so-called sampling cascade is used to study details down to the thickness of the air-blood barrier (0.5 µm, see Figure 1 of {[}\protect\hyperlink{ref-1CLJ32t5m}{50}{]} for a visualization of this cascade).
The studied volume in this case is several decimal powers less than what is studied here and still mathematically sound and accepted.

@@

It has been predicted by computational fluid dynamics simulations and by to-scale experiments that the amount and location of massless particle deposition, as well as ventilation, are depending on the size of the acini {[}\protect\hyperlink{ref-CVl41LwO}{12},\protect\hyperlink{ref-1DP2FRUSZ}{14},\protect\hyperlink{ref-RJ9GQcwl}{51}{]}.
The predictions could be confirmed at least in one case, where mice received fluorescent particles during mechanical ventilation.
In 3D-visualizations the particles were predominately detected in the proximal half of the acini {[}\protect\hyperlink{ref-19jBv0ima}{52}{]}.

Based on the above-mentioned knowledge and the data of this study, we propose the following.
In very small acini as the smallest ones, we observed at postnatal day 4 acinar flows are mainly characterized by radial streamlines {[}\protect\hyperlink{ref-eioib1TQ}{46},\protect\hyperlink{ref-afF83siN}{53}{]}.
Massless particle deposition is governed by diffusion {[}\protect\hyperlink{ref-k5fqLjiy}{54}{]}.
As the acini increase in size, convective transport becomes more and more dominant in the proximal regions of the acini.
In larger acini, the proximal region shows much higher particle retention than the distal one, because in this region the larger air-flow induces a circular current, which facilitates alveolar particle deposition.
The proximal regions may even be viewed analogously to a filter capturing the particles whereas the distal regions do not receive significant amounts {[}\protect\hyperlink{ref-CVl41LwO}{12},\protect\hyperlink{ref-1DP2FRUSZ}{14},\protect\hyperlink{ref-yHHhvOtP}{15},\protect\hyperlink{ref-1HOyVjqpM}{47},\protect\hyperlink{ref-OT4s1CSX}{48},\protect\hyperlink{ref-RJ9GQcwl}{51}{]}.

@@

We thank \texttt{manubot} {[}\protect\hyperlink{ref-YuJbg3zO}{55}{]} for providing us with the means to collaboratively write and craft this manuscript.

@@

\leavevmode\hypertarget{ref-LzalnLBH}{}%
49. Osmanagic E, Sukstanskii AL, Quirk JD, Woods JC, Pierce RA, Conradi MS, et al. Quantitative assessment of lung microstructure in healthy mice using an MR-based \textsuperscript{3} He lung morphometry technique. Journal of Applied Physiology. 2010;109: 1592--1599. doi:\href{https://doi.org/10.1152/japplphysiol.00736.2010}{10.1152/japplphysiol.00736.2010}

\leavevmode\hypertarget{ref-1CLJ32t5m}{}%
50. Weibel ER, Hsia CCW, Ochs M. How much is there really? Why stereology is essential in lung morphometry. Journal of Applied Physiology. 2007;102: 459--467. doi:\href{https://doi.org/10.1152/japplphysiol.00808.2006}{10.1152/japplphysiol.00808.2006}

\leavevmode\hypertarget{ref-RJ9GQcwl}{}%
51. Sznitman J. Convective gas transport in the pulmonary acinus: Comparing roles of convective and diffusive lengths. Journal of Biomechanics. 2009;42: 789--792. doi:\href{https://doi.org/10.1016/j.jbiomech.2008.12.022}{10.1016/j.jbiomech.2008.12.022}

\leavevmode\hypertarget{ref-19jBv0ima}{}%
52. Yang L, Feuchtinger A, Möller W, Ding Y, Kutschke D, Möller G, et al. Three-Dimensional Quantitative Co-Mapping of Pulmonary Morphology and Nanoparticle Distribution with Cellular Resolution in Nondissected Murine Lungs. ACS Nano. 2019; acsnano.8b07524. doi:\href{https://doi.org/10.1021/acsnano.8b07524}{10.1021/acsnano.8b07524}

\leavevmode\hypertarget{ref-afF83siN}{}%
53. Sznitman J, Heimsch T, Wildhaber JH, Tsuda A, Rösgen T. Respiratory Flow Phenomena and Gravitational Deposition in a Three-Dimensional Space-Filling Model of the Pulmonary Acinar Tree. Journal of Biomechanical Engineering. 2009;131: 031010. doi:\href{https://doi.org/10.1115/1.3049481}{10.1115/1.3049481}

\leavevmode\hypertarget{ref-k5fqLjiy}{}%
54. Sznitman J, Heimsch F, Altorfer D, Schittny JC, Rösgen T. Alveolar flow simulations during rhythmical breathing motion in reconstructed XTM acinar airspaces. Journal of Biomechanics. 2006;39: S441. doi:\href{https://doi.org/10.1016/s0021-9290(06)84801-3}{10.1016/s0021-9290(06)84801-3}

\leavevmode\hypertarget{ref-YuJbg3zO}{}%
55. Himmelstein DS, Rubinetti V, Slochower DR, Hu D, Malladi VS, Greene CS, et al. Open collaborative writing with Manubot. PLOS Computational Biology. 2019;15: e1007128. doi:\href{https://doi.org/10.1371/journal.pcbi.1007128}{10.1371/journal.pcbi.1007128}
\end{cslreferences}

>>>>

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681}.
It has been submited to \href{https://journals.plos.org/plosone/}{\emph{PLOS ONE}}.
The manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/5e778036e316c47b8bac4ce50086b191a7e3707a...46c4c67bf996034873f7ab33969af2db0093951a}{has been updated} since submission.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/46c4c67bf996034873f7ab33969af2db0093951a/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/46c4c67bf996034873f7ab33969af2db0093951a}{habi/acinar-analysis-manuscript@46c4c67}
on June 8, 2021.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(46c4c67)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(46c4c67)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\end{itemize}

@@

As shown in a study by Osmanagic et al.~{[}\protect\hyperlink{ref-LzalnLBH}{50}{]} the total alveolar number shows a large variability between different labs.
E.g.~the reported total number of alveoli in adult C57BL6 mice differs by a factor of approximately 8.
All the studies mentioned by Osmanagic et al.~were performed in labs well-known for their quality of stereology-based investigations of lung morphology.
Therefore, any simple bias or methodical error may not account for the reported differences.
It seems to be that the total number of alveoli is dependent on unknown small methodical differences and minute differences in how the operator actually does count.
The seemingly minuscule volume fraction of lung studied (around 1 \% or less) in this manuscript is---according to the stereological guidelines {[}\protect\hyperlink{ref-dNc8FfNn}{28}{]}---ample volume to gain mathematically correct results if the stereology is carried out correctly as we have done to the best of our knowledge.
For human lungs for example a so-called sampling cascade is used to study details down to the thickness of the air-blood barrier (0.5 µm, see Figure 1 of {[}\protect\hyperlink{ref-1CLJ32t5m}{51}{]} for a visualization of this cascade).
The studied volume in this case is several decimal powers less than what is studied here and still mathematically sound and accepted.

@@

It has been predicted by computational fluid dynamics simulations and by to-scale experiments that the amount and location of massless particle deposition, as well as ventilation, are depending on the size of the acini {[}\protect\hyperlink{ref-CVl41LwO}{12},\protect\hyperlink{ref-1DP2FRUSZ}{14},\protect\hyperlink{ref-RJ9GQcwl}{52}{]}.
The predictions could be confirmed at least in one case, where mice received fluorescent particles during mechanical ventilation.
In 3D-visualizations the particles were predominately detected in the proximal half of the acini {[}\protect\hyperlink{ref-19jBv0ima}{53}{]}.

Based on the above-mentioned knowledge and the data of this study, we propose the following.
In very small acini as the smallest ones, we observed at postnatal day 4 acinar flows are mainly characterized by radial streamlines {[}\protect\hyperlink{ref-eioib1TQ}{46},\protect\hyperlink{ref-afF83siN}{54}{]}.
Massless particle deposition is governed by diffusion {[}\protect\hyperlink{ref-k5fqLjiy}{55}{]}.
As the acini increase in size, convective transport becomes more and more dominant in the proximal regions of the acini.
In larger acini, the proximal region shows much higher particle retention than the distal one, because in this region the larger air-flow induces a circular current, which facilitates alveolar particle deposition.
The proximal regions may even be viewed analogously to a filter capturing the particles whereas the distal regions do not receive significant amounts {[}\protect\hyperlink{ref-CVl41LwO}{12},\protect\hyperlink{ref-1DP2FRUSZ}{14},\protect\hyperlink{ref-yHHhvOtP}{15},\protect\hyperlink{ref-1HOyVjqpM}{47},\protect\hyperlink{ref-OT4s1CSX}{48},\protect\hyperlink{ref-RJ9GQcwl}{52}{]}.

@@

We thank \texttt{manubot} {[}\protect\hyperlink{ref-YuJbg3zO}{56}{]} for providing us with the means to collaboratively write and craft this manuscript.

@@

\leavevmode\hypertarget{ref-VJYqPtdT}{}%
49. Das GK, Anderson DS, Wallis CD, Carratt SA, Kennedy IM, Van Winkle LS. Novel multi-functional europium-doped gadolinium oxide nanoparticle aerosols facilitate the study of deposition in the developing rat lung. Nanoscale. 2016;8: 11518--11530. doi:\href{https://doi.org/10.1039/c6nr00897f}{10.1039/c6nr00897f}

\leavevmode\hypertarget{ref-LzalnLBH}{}%
50. Osmanagic E, Sukstanskii AL, Quirk JD, Woods JC, Pierce RA, Conradi MS, et al. Quantitative assessment of lung microstructure in healthy mice using an MR-based \textsuperscript{3} He lung morphometry technique. Journal of Applied Physiology. 2010;109: 1592--1599. doi:\href{https://doi.org/10.1152/japplphysiol.00736.2010}{10.1152/japplphysiol.00736.2010}

\leavevmode\hypertarget{ref-1CLJ32t5m}{}%
51. Weibel ER, Hsia CCW, Ochs M. How much is there really? Why stereology is essential in lung morphometry. Journal of Applied Physiology. 2007;102: 459--467. doi:\href{https://doi.org/10.1152/japplphysiol.00808.2006}{10.1152/japplphysiol.00808.2006}

\leavevmode\hypertarget{ref-RJ9GQcwl}{}%
52. Sznitman J. Convective gas transport in the pulmonary acinus: Comparing roles of convective and diffusive lengths. Journal of Biomechanics. 2009;42: 789--792. doi:\href{https://doi.org/10.1016/j.jbiomech.2008.12.022}{10.1016/j.jbiomech.2008.12.022}

\leavevmode\hypertarget{ref-19jBv0ima}{}%
53. Yang L, Feuchtinger A, Möller W, Ding Y, Kutschke D, Möller G, et al. Three-Dimensional Quantitative Co-Mapping of Pulmonary Morphology and Nanoparticle Distribution with Cellular Resolution in Nondissected Murine Lungs. ACS Nano. 2019; acsnano.8b07524. doi:\href{https://doi.org/10.1021/acsnano.8b07524}{10.1021/acsnano.8b07524}

\leavevmode\hypertarget{ref-afF83siN}{}%
54. Sznitman J, Heimsch T, Wildhaber JH, Tsuda A, Rösgen T. Respiratory Flow Phenomena and Gravitational Deposition in a Three-Dimensional Space-Filling Model of the Pulmonary Acinar Tree. Journal of Biomechanical Engineering. 2009;131: 031010. doi:\href{https://doi.org/10.1115/1.3049481}{10.1115/1.3049481}

\leavevmode\hypertarget{ref-k5fqLjiy}{}%
55. Sznitman J, Heimsch F, Altorfer D, Schittny JC, Rösgen T. Alveolar flow simulations during rhythmical breathing motion in reconstructed XTM acinar airspaces. Journal of Biomechanics. 2006;39: S441. doi:\href{https://doi.org/10.1016/s0021-9290(06)84801-3}{10.1016/s0021-9290(06)84801-3}

\leavevmode\hypertarget{ref-YuJbg3zO}{}%
56. Himmelstein DS, Rubinetti V, Slochower DR, Hu D, Malladi VS, Greene CS, et al. Open collaborative writing with Manubot. PLOS Computational Biology. 2019;15: e1007128. doi:\href{https://doi.org/10.1371/journal.pcbi.1007128}{10.1371/journal.pcbi.1007128}
\end{cslreferences}
